---
title: "Exploratory Data Analysis Of laptops dataset"
author: "Usama Ali"
date: "2023-09-30"
output: html_document
---

### Objective
- The objective of this project was to perform an exploratory data analysis of a laptop dataset using R and ggplot2 for data visualization. The analysis aimed to extract valuable insights from the dataset and provide clear visualizations to enhance understanding.

### Dataset used:
- About: This dataset provides a comprehensive collection of information on various laptops, enabling a detailed analysis of their specifications and pricing. It encompasses a wide range of laptops, encompassing diverse brands, models, and configurations, making it a valuable resource for researchers, data analysts, and machine learning enthusiasts interested in the laptop industry.

- Source: Kaggle.com

- License: CCO Public Domain

- Uploaded By: Juan Merino 

#### Importing the libraries
```{r import, message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library(markdown)
```

#### Importing Dataset
```{r loading datset, message=FALSE, warning=FALSE}
laptops <- read.csv('laptops.csv')
```

#### Checking for duplicates & removing
```{r duplicates, message=FALSE, warning=FALSE, include=FALSE}

#checking for duplicates
#duplicated(laptops)

#getting rid of duplicated rows
laptops <- distinct(laptops)
```

#### Taking a look and cleaning
```{r cleaning}

#viewing the first rows and structure
glimpse(laptops)

#basic descriptive stats for our dataset & Na values
skim_without_charts(laptops)

#removing missing values
laptops <- na.omit(laptops)
```

### Plot Analysis

#### Price Distribution:
```{r histogram}
#histogram for price distribution
ggplot(laptops, aes(x = Final.Price)) +
  geom_histogram(binwidth = 200, fill="skyblue", color = 'black')+
  labs(title = 'Distribution of Laptop Prices',
       x = 'Final Price', y = 'Frequency')
```

- The histogram displays a right-skewed distribution. Three prominent peaks are observed, with frequencies exceeding 350 for one peak.

- As prices increase beyond $300, the frequency gradually declines until the $5000 price range. Between $5000 and $6000, the distribution becomes less visible.
Laptops priced above $6000 are exceedingly rare in the dataset.


#### Brand Analysis - Average Price:
```{r barchart}
#brand analysis, avg_price for each brand bar chart
brand_avg_prices <- laptops %>% 
  group_by(Brand) %>% 
  summarise(avg_price = mean(Final.Price))
ggplot(brand_avg_prices, aes(x = Brand, y = avg_price, 
                             fill = avg_price))+
  geom_bar(stat = 'identity') +
  labs(x = 'Brand', y = 'Avergae Price', 
       title = 'Average Laptop Prices by Brand') +
  theme(axis.text.x = element_text(angle = 45))
```

- The bar chart presents the average prices of laptops by brand.

- Razer stands out as the highest-priced brand, with laptops peaking at around $3500 in average price.

- Millenium and MSI are the second and third highest-priced brands, with laptops sitting near the $2000 price mark on average.

- Popular brands like Apple, Microsoft, and Gigabyte have laptops priced at nearly $1500 on average.

#### RAM & Price Relationship:
```{r scatter plot}
#RAM & Price relationship via scatter plot
ggplot(laptops, aes(x = RAM, y = Final.Price))+
  geom_point()+
  labs(title = 'Scattor Plot of RAM vs Laptop Price',
       x = "RAM", y = 'Final Price')
```

- The scatter plot reveals the relationship between laptop RAM (in GB) and prices.

- For laptops with 4GB RAM, a cluster is observed in the $200-$800 price range.Laptops with 8GB RAM cluster between $200 and $2000, with fewer above $2000. The 16GB RAM category is dense, spanning $200-$4000, with fewer observations beyond $3000.

-64GB RAM laptops are sparsely represented.
An isolated 120GB RAM laptop is priced around $6000, suggesting uniqueness.

#### Storage Type Distribution:
```{r piechart}
#distribution of storage types using pie chart
storage_distribution <- table(laptops$Storage.type)
pie(storage_distribution, labels = names(storage_distribution),
    main = "Distribution of Storage Types")
```

- The pie chart illustrates the predominance of SSD storage types among laptops.

- SSD storage types represent the majority of laptops in the dataset, while eMMC storage types constitute a very small proportion.


#### Touchscreen vs. Non-Touchscreen Prices:
```{r box-plot}
#5.box-plot for avg prices of touchscreen vs non-touchscreen
ggplot(laptops, aes(x = Touch, y = Final.Price))+
  geom_boxplot()+
  labs(x = 'Touchscreen', y = 'Final Price',
       title = 'Box Plot of Prices by Touchscreen Availability')
```

- The box plot compares laptop prices between two categories: "Yes" (touchscreen) and "No" (non-touchscreen).

- The median for the "No" category sits at approximately $1000, while for the "Yes" category, it's slightly above $1500.

- In the "No" category (non-touchscreen), the majority of data points are clustered above the 3rd quartile, ranging from approximately $3000 to $6000. There is a notable outlier priced above $7000.

- The median is positioned between the 1st and 2nd quartiles but is also closer to the center.

- In the "Yes" category (touchscreen), the pattern is similar, with most data points concentrated above the 3rd quartile, at around $4500.

- The median leans toward the 3rd quartile rather than the 1st quartile, indicating differences in price distribution between the two categories.

#### CPU Analysis - Most Common CPU
```{r bar}
#CPU analysis, most common CPU barplot
cpu_freq <- table(laptops$CPU)
common_cpu <- names(which.max(cpu_freq))
barplot(cpu_freq,
        main = 'CPU Frequency in Laptops',
        ylab = "Frequency", col = 'skyblue',
        las= 2)
```

- The bar chart displays the frequency of different CPU types among laptops.

- Intel Core i7 CPUs are the most prevalent, with a frequency of approximately 700. Intel Core i5 CPUs follow as the second most common, with a frequency of around 500.

- AMD Ryzen 7 CPUs are the third highest in frequency, at approximately 150.

- Ryzen 5 CPUs are the fourth highest in frequency.

### Recommendations:

- Consider exploring the unique features and specifications of laptops from brands like Razer, Millenium, and MSI, as they offer higher-priced options with potential performance benefits.

- For consumers seeking laptops within the $200 to $800 price range, focus on laptops with 4GB and 8GB RAM, which are prevalent and offer good value.

- Manufacturers should take note of the popularity of Intel Core i7 and Core i5 CPUs and potentially offer a wider range of laptops featuring these processors.
